---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
 

## Loading and preprocessing the data
Load activity.csv and replace NA with 0, check for content

```{r}
a <- read.csv("activity.csv" ) 
a[is.na(a)] <- 0
```

## What is mean total number of steps taken per day?
Aggregate data by date, compute total number of steps for each day, take a mean of that list.
```{r}
agg <- aggregate(a$steps,by=list(date=a$date),sum)
head(agg)
means <- round(mean(agg[,2]))
medians <- as.integer(median(agg[,2]))
```

Plot the histogram of number of steps per day
```{r}
hist(agg[,2],xlab='steps',ylab='days', main='Steps per day')
```

Mean total number of steps per day = `r means`  
Median total number of steps per day = `r medians`

## What is the average daily activity pattern?
```{r}
agint <- aggregate(a$steps,by=list(interval=a$interval),mean)
head(agint)
tail(agint)
plot(agint$interval,agint$x,type="l",xlab="time",ylab="steps")
moststep <- agint$interval[which.max(agint$x)]
```

Most active interval is `r moststep`.

## Imputing missing values, using previously computed mean for that interval
```{r}
b <- read.csv("activity.csv" ) 
m <- is.na(b$steps)
mc <- nrow(b)
tm <- sum(m)
for(i in (1:mc)) 
  {
    if(is.na(b$steps[i])) {
      b$steps[i] <- agint$x[match(b$interval[i],agint$interval)]
      }
  }
```
Total missing values: `r tm` out of `r mc`.

```{r}
## histogram and mean/median for new dataset
aggb <- aggregate(b$steps,by=list(date=b$date),sum)
head(aggb)
meansb <- as.integer(round(mean(aggb[,2])))
mediansb <- as.integer(round(median(aggb[,2])))

hist(aggb[,2],xlab='steps',ylab='days', main='Steps per day (with imputed data)')
```

After imputing the data:  
Mean total number of steps per day = `r meansb`.
Median total number of steps per day = `r mediansb`.

We see that median remained the same (half of imputed values should be below median, half above - makes sense) while mean has grown (also makes sense as we use computed mean numbers instead of 0s).

## Are there differences in activity patterns between weekdays and weekends?
Adding new factor
```{r}
a$DayOfWeekType <- ifelse(weekdays(as.Date(a$date)) %in% c('Saturday','Sunday'), 'Weekend', 'Weekday')
agdayint <- aggregate(a$steps,by=list(interval=a$interval,day=a$DayOfWeekType),mean)
head(agdayint)
library(ggplot2)
qplot(interval,x,data=agdayint,facets=day~.,geom='line',ylab='number of steps')
```
We can see a difference between weekdays and weekends - as people make fewer steps during weekend mornings - we can guess people sleep in longer on the weekends :)
